---
title: "Global temperature predictions accuracy"
author: "Gonçalo Nogueira"
date: "September 17, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
```

## IPCC assesment reports

The Intergovernmental Panel on Climate Change (IPCC) produces assessments reports on the current
state of the climate change. The first assessment report was produced in 1990 and contained a
prediction of the global temperature change.

We will see how the predictions match up with reality based on data on measured global temperature
provided by NASA.

## Predictions of the 1st Assessment Report

The IPCC predicted a rise of temperatures of 0.7 to 1.5 ºC from 1990 to 2030.
So the various scenarios are modeled by the following linear relationships:

$$T_i = T_{1990} + b(i - 1990)$$

$T_i$ - predicted global temperature in year $i$

$T_{1990}$ - observed global temperature in 1990

$b$ - temperature change by year; depends on the scenario being considered

## Scenarios being evaluated

**Low scenario** - 0.7 ºC increase in 2030 (0.0175 ºC per year)

**Intermediate scenario** - 1.1 ºC increase in 2030 (0.0275 ºC per year)

**High scenario** - 1.5 ºC increase in 2030 (0.0375 ºC per year)

## Visualization of the predictions

```{r}
temp <- read.table("nasa-global-temp.txt", header = FALSE,
                   col.names = c("year", "temperature", "temperature.5.year.mean"))

temp$low.scenario <- rep(NA, nrow(temp))
temp$intermediate.scenario <- rep(NA, nrow(temp))
temp$low.scenario <- rep(NA, nrow(temp))

temp.prediction <- function(year, b) {
  temp[temp$year == 1990, "temperature"] + b * (year - 1990)
}

temp[temp$year >= 1990, "low.scenario"] <- sapply(temp[temp$year >= 1990, "year"],
                                                  temp.prediction, 0.0175)

temp[temp$year >= 1990, "intermediate.scenario"] <- sapply(temp[temp$year >= 1990, "year"],
                                                      temp.prediction, 0.0275)

temp[temp$year >= 1990, "high.scenario"] <- sapply(temp[temp$year >= 1990, "year"],
                                                      temp.prediction, 0.0375)

plot_ly(temp, x = ~year, y = ~temperature, name = "Measured global temperature",
        type = "scatter", mode = "lines") %>%
  add_trace(y = ~low.scenario, name = "Low scenario", mode = "lines") %>%
  add_trace(y = ~intermediate.scenario, name = "Intermediate scenario", mode = "lines") %>%
  add_trace(y = ~high.scenario, name = "High scenario", mode = "lines") %>%
  layout(title = "IPCC 1st AR predictions VS measured global temperature",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Relative temperature (ºC)"),
         hovermode = "closest")
```
